---
title: "SRDP Control Variables Codebook"
output:
  pdf_document
header-includes: |
  \usepackage{fancyhdr}
  \pagestyle{fancy}
  \fancypagestyle{plain}{\pagestyle{fancy}}
  \fancyhead[LE, RO]{Strategies of Resistance Data Project}
  \fancyhead[RE, LO]{`r format(Sys.time(), '%B %d, %Y')`}
fontsize: 12pt
font-family: Calibri
geometry: margin=0.8in
spacing: double
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r libraries}
library(tidyverse)
library(rio)
library(here)
```

```{r datasets}
# country level
country_level <- import(here("data", "country_level.csv"))
country_list <- country_level %>% distinct(country) %>% pull()

# group level
group_level <- import(here("data", "group_level.csv"))
```

# Scope

SRDP covers all countries with organizations actively pursuing self-determination claims from 1960 to 2003 included in the CIDCM Peace and Conflict report. This includes `r country_level %>% distinct(country) %>% nrow()` countries: `r country_level %>% distinct(country) %>% pull()`.

# Country-level data

## Region

**region**

*The geographic region of the country.*

Unit: Region.

Source: World Bank

## Population

**population**

*Total population is based on the de facto definition of population, which counts all residents regardless of legal status or citizenship. The values shown are midyear estimates.*

Unit: Total population

Source: [United Nations Department of Economic and Social Affairs](https://population.un.org/wpp/Download/Standard/Population/)

Supplementary source: [World Bank](https://data.worldbank.org/indicator/SP.POP.TOTL)

Year range: `r country_level %>% filter(!is.na(population)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(population)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% group_by(country) %>% filter(is.na(population)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## GDP

**gdp**

*GDP at purchaser's prices is the sum of gross value added by all resident producers in the economy plus any product taxes and minus any subsidies not included in the value of the products. It is calculated without making deductions for depreciation of fabricated assets or for depletion and degradation of natural resources. Data are in current U.S. dollars. Dollar figures for GDP are converted from domestic currencies using single year official exchange rates. For a few countries where the official exchange rate does not reflect the rate effectively applied to actual foreign exchange transactions, an alternative conversion factor is used.*

Unit: Current USD

Source: [World Bank](https://data.worldbank.org/indicator/NY.GDP.MKTP.CD)

Year range: `r country_level %>% filter(!is.na(gdp)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(gdp)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% group_by(country) %>% filter(is.na(gdp)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Military expenditure

**milex**

*Military expenditure in local currency at current prices is presented according to both the financial year of each country and according to calendar year, calculated on the assumption that, where financial years do not correspond to calendar years, spending is distributed evenly through the year. Figures in current US \$ are presented according to calendar year.*

Unit: Current USD

Source: [SIRPI](https://www.sipri.org/databases/milex)

Year range: `r country_level %>% filter(!is.na(milex)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(milex)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% group_by(country) %>% filter(is.na(milex)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Unified Democracy Scores

**uds**

*Unified Democracy Scores (UDS) are a set of measures that leverage the efforts of a variety of experts to provide a composite scale of democracy, accompanied by estimates of measurement uncertainty.*

Unit: Unified democracy score posterior mean

Source: [UDS](http://www.unified-democracy-scores.net/)

Year range: `r country_level %>% filter(!is.na(uds)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(uds)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% filter(year < 2013) %>% group_by(country) %>% filter(is.na(uds)) %>% distinct(country) %>% pull()`

Update schedule: No longer updated

## Freedom House political rights

**fh_political_rights**

*A composite measure of the political rights citizens of a country enjoy. For a more complete definition, see [Freedom House's methodology](https://freedomhouse.org/reports/freedom-world/freedom-world-research-methodology).*

Unit: See above.

Source: [Freedom House](https://freedomhouse.org/report/freedom-world/2021/democracy-under-siege)

Year range: `r country_level %>% filter(!is.na(fh_political_rights)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(fh_political_rights)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% filter(year > 1971) %>% group_by(country) %>% filter(is.na(fh_political_rights)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Freedom House civil liberties

**fh_civil_liberties**

*A composite measure of the civil liberties citizens of a country enjoy. For a more complete definition, see [Freedom House's methodology](https://freedomhouse.org/reports/freedom-world/freedom-world-research-methodology).*

Unit: See above.

Source: [Freedom House](https://freedomhouse.org/report/freedom-world/2021/democracy-under-siege)

Year range: `r country_level %>% filter(!is.na(fh_civil_liberties)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(fh_civil_liberties)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% filter(year > 1971) %>% group_by(country) %>% filter(is.na(fh_civil_liberties)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Freedom House status

**fh_status**

*A categorical measure of the political rights and civil liberties citizens of a country enjoy. For a more complete definition, see [Freedom House's methodology](https://freedomhouse.org/reports/freedom-world/freedom-world-research-methodology).*

Unit: See above.

Source: [Freedom House](https://freedomhouse.org/report/freedom-world/2021/democracy-under-siege)

Year range: `r country_level %>% filter(!is.na(fh_status)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(fh_status)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% filter(year > 1971) %>% group_by(country) %>% filter(is.na(fh_status)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Polity democracy score

**polity_democracy**

*A composite measure of democracy in a country. For a more complete definition, see the [PolityV methodology](https://www.systemicpeace.org/inscr/p4manualv2016.pdf).*

Unit: Score from 0 to 10. A higher score indicates higher levels of democracy.

Source: [PolityV](http://www.systemicpeace.org/inscrdata.html)

Year range: `r country_level %>% filter(!is.na(polity_democracy)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(polity_democracy)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% group_by(country) %>% filter(is.na(polity_democracy)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Polity autocracy score

**polity_autocracy**

*A composite measure of autocracy in a country. For a more complete definition, see the [PolityV methodology](https://www.systemicpeace.org/inscr/p4manualv2016.pdf).*

Unit: Score from 0 to 10. A higher score indicates higher levels of autocracy.

Source: [PolityV](http://www.systemicpeace.org/inscrdata.html)

Year range: `r country_level %>% filter(!is.na(polity_autocracy)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(polity_autocracy)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% group_by(country) %>% filter(is.na(polity_autocracy)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Polity total score

**polity_total**

*A composite measure of the net level of democracy and autocracy in a country. Caluculated by subtracting the country's autocracy score from its democracy score. For a more complete definition, see the [PolityV methodology](https://www.systemicpeace.org/inscr/p4manualv2016.pdf).*

Unit: See above. Ranges from -10 (full autocracy) to 10 (full democracy).

Source: [PolityV](http://www.systemicpeace.org/inscrdata.html)

Year range: `r country_level %>% filter(!is.na(polity_total)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(polity_total)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% group_by(country) %>% filter(is.na(polity_total)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Checks and balances

**checks**

*Measure of the degree to which a country's executive and legislature have checks and balances on their power. Equal to one if the country has:*

1.  *No legislature or executive,*

2.  *Unelected legislature or executive,*

3.  *Elected, one candidate,*

4.  *One party, multiple candidates,*

5.  *Multiple legal parties but only one party won seats,*

6.  *Multiple parties did win seats but the largest party received more than 75% of the seats.*

*Countries where legislatures are not competitively elected are considered countries where only the executive wields a check. Incremented by one if there is a chief executive. Incremented by one if the chief executive is competitively elected. Incremented by one if the opposition controls the legislature.*

*In presidential systems, increases by one for (1) each chamber of the legislature unless the president's party has a majority in the lower house and a closed list system is in effect (implying stronger presidential control of his/her party, and therefore of the legislature), and (2) for each party coded as allied with the president's party and which has an ideological (left-right-center) orientation closer to that of the main opposition party than to that of the president's party.*

*In parliamentary systems, increases by one for (1) every party in the government coalition as long as the parties are needed to maintain a majority, and (2) every party in the government coalition that has a position on economic issues (right-left-center) closer to the largest opposition party than to the party of the executive. In parliamentary systems, the prime minister's party is not counted as a check if there is a closed rule in place: the prime minister is presumed in this case to control the party fully.*

Unit: See above.

Source: [Database of Political Institutions](https://data.iadb.org/DataCatalog/Dataset#DataCatalogID=11319/29752)

Year range: `r country_level %>% filter(!is.na(checks)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(checks)) %>% summarise(max(year))`

Countries with incomplete data: `r country_level %>% group_by(country) %>% filter(is.na(checks)) %>% distinct(country) %>% pull()`

Update schedule: Annual

## Civil war incidence

**civil_war**

*Binary variable describing indicating the occurrence of a civil war leading to at least twenty-five battle-deaths in a calendar year. This measure includes intrastate and internationalized intrastate wars (according to the UCDP/PRIO definition).*

Unit: Binary.

Source [UCDP/PRIO Armed Conflict Dataset version 21.1](https://ucdp.uu.se/downloads/index.html#armedconflict)

Year range: `r country_level %>% filter(!is.na(civil_war)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(civil_war)) %>% summarise(max(year))`

Countries with incomplete data: None.

Update schedule: Annual

## Civil war onset

**civil_war_onset**

*Binary variable indicating the outbreak of civil war leading to at least twenty-five battle-deaths in a calendar year preceded by at least two years of peace. This measure includes intrastate and internationalized intrastate wars (according to the UCDP/PRIO definition).*

Unit: Binary.

Source [UCDP/PRIO Armed Conflict Dataset version 21.1](https://ucdp.uu.se/downloads/index.html#armedconflict)

Year range: `r country_level %>% filter(!is.na(civil_war_onset)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(civil_war_onset)) %>% summarise(max(year))`

Countries with incomplete data: None.

Update schedule: Annual

## Civil war in previous year

**civil_war_prev_yr**

*Binary variable indicating civil war in the state in the previous calendar year (twenty-five battle-deaths threshold). This measure includes intrastate and internationalized intrastate wars (according to the UCDP/PRIO definition).*

Unit: Binary.

Source [UCDP/PRIO Armed Conflict Dataset version 21.1](https://ucdp.uu.se/downloads/index.html#armedconflict)

Year range: `r country_level %>% filter(!is.na(civil_war_prev_yr)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(civil_war_prev_yr)) %>% summarise(max(year))`

Countries with incomplete data: None.

Update schedule: Annual

## Elections

**election**

*The number of elections held in the country during the calendar year.*

Source: [International Foundation for Electoral Systems](https://www.electionguide.org/)

Year range: `r country_level %>% filter(!is.na(election)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(election)) %>% summarise(max(year))`

Countries with incomplete data: None.

Update schedule: Weekly.

## Referendums

**referendum**

*The number of referendums held in the country during the calendar year.*

Source: [International Foundation for Electoral Systems](https://www.electionguide.org/)

Year range: `r country_level %>% filter(!is.na(referendum)) %>% summarise(min(year))` to `r country_level %>% filter(!is.na(referendum)) %>% summarise(max(year))`

Countries with incomplete data: None.

Update schedule: Weekly.

# Group-level data

## Relative size

**relative_size**

*The groups' relative sizes as a share of the country's total population.*

Source: [Ethnic Power Relations](https://icr.ethz.ch/data/epr/core/) and [United Nations Department of Economic and Social Affairs](https://population.un.org/wpp/Download/Standard/Population/)

Supplementary sources: [Minorities at Risk](http://www.mar.umd.edu/mar_data.asp)

Year range: `r group_level %>% filter(!is.na(relative_size)) %>% summarise(min(year))` to `r group_level %>% filter(!is.na(relative_size)) %>% summarise(max(year))`

Groups with incomplete data: `r group_level %>% group_by(groupname) %>% filter(is.na(relative_size)) %>% distinct(groupname) %>% pull()`

Missing data: Data are first collected from the EPR. Next, data were collected from Minorities at Risk. MAR only provides data from 2004 to 2006. SRDP uses the average group size across 2004 to 2006 to calculate each group's size relative to the country's population (drawn from the `population` variable described above). To maintain as much consistency as possible, SRDP does not use the relative group population size provided by MAR for 2004 to 2006. Rather, we calculate groups' relative size in 2004 to 2006 using the group's absolute size and population data.

Update schedule: Annual

## Size

**size**

*The estimated total population of the group.*

Source: [Ethnic Power Relations](https://icr.ethz.ch/data/epr/core/)

Supplementary sources: [Minorities at Risk](http://www.mar.umd.edu/mar_data.asp)

Year range: `r group_level %>% filter(!is.na(relative_size)) %>% summarise(min(year))` to `r group_level %>% filter(!is.na(relative_size)) %>% summarise(max(year))`

Groups with incomplete data: `r group_level %>% group_by(groupname) %>% filter(is.na(relative_size)) %>% distinct(groupname) %>% pull()`

Missing data: Data are first collected from the EPR. Next, data were collected from Minorities at Risk. MAR only provides data from 2004 to 2006. SRDP uses the average group size across 2004 to 2006 to provide values for each missing year.

Update schedule: Annual
